Lyman-$\alpha$ photons through rotating outflows

### 2018-07-17 paper/paper.tex
fixing running head
some fixes

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title[Lyman-α photons through rotating outflows]{Lyman-$\alpha$ photons through rotating outflows}
\author[M.C. Remolina-Gutierrez \& J.E. Forero-Romero]{
  Maria Camila Remolina-Guti\'errez$^{1}$
  \thanks{mc.remolina197@uniandes.edu.co} \&
  Jaime E. Forero-Romero $^{1}$
  \thanks{je.forero@uniandes.edu.co}\\
  %%
  $^{1}$ Departamento de F\'isica, Universidad de los Andes, Cra. 1
  No. 18A-10 Edificio Ip, CP 111711, Bogot\'a, Colombia \\
}

@@

Recent advances in instrumentation have revealed the presence of gas
rotation on vastly different physical scales.
For instance, spatially resolved spectra on compact dwarf galaxies
have measured clear signs of gas showing pure rotation kinematics
\citep{2015A&A...577A..21C,2017A&A...600A.125C} and the recent mapping of
high redshift circumgalactic regions have also revealed kinematic
evidence for large scale rotation \citep{2018MNRAS.473.3907A}.
These systems with star formation, neutral gas and low dust
contents naturally show a \lya emission line \citep{PartridgePeebles}
motivating the observational work to phenomenologically link
H$\alpha$ kinematics to \lya spectra \citep[e.g.][]{Herenz2016}.

@@

However, the gas dynamics in Lyman Alpha Emitter (LAE) galaxies are
more complex than pure rotation.
The case for outflows as traced in the \lya line is solidly
established.  
In many observations the \lya line profile has a single peak
redwards from the line's center, in other cases there is a double peak
but the peak on the red side is stronger
\citep[e.g.][]{2010ApJ...717..289S,Erb14,Trainor16}.   
These features have been explained as the consequence of multiple
\lya photon scatterings through a homogeneous outflowing shell of
neutral Hydrogen
\citep{2006A&A...460..397V,Orsi12,2012ApJ...751...29Y,2015ApJ...812..123G}.

@@

As input parameters we use $\tauh=\{10^5, 10^6, 10^7\}$,
$\vout=\{5,25,50\}\;\kms$ and $\vrot=\{0,50,100\}\;\kms$, for a total of
$27$ models with all the possible parameter combinations.
The range of values have some overlap with the expectations from a
galaxy with a total neutral hydrogen mass of $10^8$-$10^9$
$M_{\odot}$.

@@

\begin{equation}
x' = x + \frac{\vec{v}_{\rm rot}\cdot \hatk}{\vth}
\label{eq:shift_x}
\end{equation}
%
where $x'$ is the photon's new adimensional frequency, $x$ is the photon's
frequency after being processed only by the outflow, \vrot is the 
rotational velocity at the point of escape of the photon, \hatk is
the photon's direction of propagation and \vth is the thermal
velocity of the sphere.
We also define the viewing angle, $\theta$, as the angle between the
rotation axis and the line of sight of a potential observer.

@@

\subsection{Quantitative trends}
\label{sec:quantitative}

>>>>

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title[Lyman-$\alpha$ photons through rotating outflows]{Lyman-$\alpha$ photons through rotating outflows}
\author[M.C. Remolina-Gutierrez \& J.E. Forero-Romero]{
  Maria Camila Remolina-Guti\'errez$^{1}$
  \thanks{mc.remolina197@uniandes.edu.co} \&
  Jaime E. Forero-Romero $^{1}$
  \thanks{je.forero@uniandes.edu.co}\\
  %%
  $^{1}$ Departamento de F\'isica, Universidad de los Andes, Cra. 1
  No. 18A-10 Edificio Ip, CP 111711, Bogot\'a, Colombia \\
}

@@

Recent advances in instrumentation have revealed the presence of gas
rotation on vastly different physical scales.
For instance, spatially resolved spectra on compact dwarf galaxies
have measured clear signs of gas showing pure rotation kinematics
\citep{2015A&A...577A..21C,2017A&A...600A.125C} and the recent mapping of
high redshift circumgalactic regions have also revealed kinematic
evidence for large scale rotation \citep{2018MNRAS.473.3907A}.
These systems with star formation, neutral gas and low dust
contents naturally produce a \lya emission line \citep{PartridgePeebles}
motivating the observational work to phenomenologically link
H$\alpha$ kinematics to \lya spectra \citep[e.g.][]{Herenz2016}.

@@

However, the gas dynamics in Lyman Alpha Emitter (LAE) galaxies are
more complex than pure rotation.
The case for outflows as traced in the \lya line is solidly
established.  
In many observations the \lya line profile has a single peak
redwards from the line's center, in other cases there is a double peak
but the peak on the red side is stronger
\citep[e.g.][]{2010ApJ...717..289S,Erb14,Trainor16}.   
These features have been explained as the consequence of multiple
\lya photon scatterings through a homogeneous outflowing shell of
neutral Hydrogen
\citep{2006A&A...460..397V,Orsi12,2015ApJ...812..123G}.

@@

As input parameters we use $\tauh=\{10^5, 10^6, 10^7\}$,
$\vout=\{5,25,50\}\;\kms$ and $\vrot=\{0,50,100\}\;\kms$, for a total of
$27$ models with all the possible parameter combinations.
The range of values have some overlap with the expectations from a
galaxy with a total neutral hydrogen mass of $10^8$-$10^9$
$M_{\odot}$. We also define the viewing angle, $\theta$, as the angle
between the rotation axis and the line of sight of a potential observer.

@@

\begin{equation}
x' = x + \frac{\vec{v}_{\rm rot}\cdot \hatk}{\vth}
\label{eq:shift_x}
\end{equation}
%
where $x'$ is the photon's new adimensional frequency, $x$ is the photon's
frequency after being processed only by the outflow, \vrot is the 
rotational velocity at the point of escape of the photon, \hatk is
the photon's direction of propagation and \vth is the thermal
velocity of the sphere. In this study, the latter is set to a fixed value
of \vth $= 12.86$ \kms, which corresponds to a temperature of $T=10^4$ K. The reason of setting it fixed is so that the comparison between the simulations and the semianalitical model is clearer, because there is one less parameter space to explore. So, we set \vth to a constant and consistent value throughout the analysis. However we are aware of the smoothing effect it has over the Doppler shift, as seen in Eq. \ref{eq:shift_x}, and it should be an explicitly free parameter if this model is used to fit observations.

@@

\subsection{Quantitative Trends}
\label{sec:quantitative}
